---
title: "gtsummary"
search: true
---

```{r warning = F, message = F}
library(tidyverse)
library(gtsummary)
```

## Links

+ [gtsummary](https://www.danieldsjoberg.com/gtsummary/index.html)
+ [gt](https://gt.rstudio.com/)

## Basic Usage

### Example 1 

```{r}
trial %>% tbl_summary(include = c(trt, age, grade))
```

> For continuous variable, present `median (Q1, Q3)`

| Argument | Description |
|:---|:---|
| `label` | specify the variable labels printed in table |
| `type` | specify the variable type (e.g. continuous, categorical, etc.) |
| `statistic` | change the summary statistics presented |
| `digits` | number of digits the summary statistics will be rounded to |
| `missing` | whether to display a row with the number of missing observations |
| `missing_text` | text label for the missing number row |
| `missing_stat` | statistic(s) to show on the missing row |
| `sort` | change the sorting of categorical levels by frequency |
| `percent` | print column, row, or cell percentages |
| `include` | list of variables to include in summary table |

## Advanced Usage

### Example 2: Modify **`tbl_summary()`** arguments 

```{r}
trial |>
  tbl_summary(
    by = trt,
    include = c(age, grade),
    statistic = list(
      all_continuous() ~ "{mean} ({sd})",
      all_categorical() ~ "{n} / {N} ({p}%)"
    ),
    digits = all_continuous() ~ 2,
    label = list(grade = "Tumor Grade"),
    missing_text = "(Missing)"
  )
```

| Function | Description |
|:---|:---|
| [`add_p()`](https://www.danieldsjoberg.com/gtsummary/reference/add_p.html) | add p-values to the output comparing values across groups |
| [`add_overall()`](https://www.danieldsjoberg.com/gtsummary/reference/add_overall.html) | add a column with overall summary statistics |
| [`add_n()`](https://www.danieldsjoberg.com/gtsummary/reference/add_n.html) | add a column with N (or N missing) for each variable |
| [`add_difference()`](https://www.danieldsjoberg.com/gtsummary/reference/add_difference.html) | add column for difference between two group, confidence interval, and p-value |
| [`add_stat_label()`](https://www.danieldsjoberg.com/gtsummary/reference/add_stat_label.html) | add label for the summary statistics shown in each row |
| [`add_stat()`](https://www.danieldsjoberg.com/gtsummary/reference/add_stat.html) | generic function to add a column with user-defined values |
| [`add_q()`](https://www.danieldsjoberg.com/gtsummary/reference/add_q.html) | add a column of q values to control for multiple comparisons |

### Example 3: Modify Table 

```{r}
trial |>
  tbl_summary(by = trt, includ = c(age, grade)) |>
  add_p(pvalue_fun = label_style_pvalue(digits = 2)) |>
  add_overall() |>
  add_n() |>
  modify_header(label ~ "**Variable**") |>
  modify_spanning_header(c("stat_1", "stat_2") ~ "**Treatment Received**") |>
  modify_footnote_header("Median (IQR) or Frequency (%)", columns = all_stat_cols()) |>
  modify_caption("**Table 1. Patient Characteristics**") |>
  bold_labels()
```

| Function | Description |
|:---|:---|
| [`modify_header()`](https://www.danieldsjoberg.com/gtsummary/reference/modify.html) | update column headers |
| [`modify_footnote_header()`](https://www.danieldsjoberg.com/gtsummary/reference/modify_footnote2.html) | update column header footnote |
| [`modify_footnote_body()`](https://www.danieldsjoberg.com/gtsummary/reference/modify_footnote2.html) | update table body footnote |
| [`modify_spanning_header()`](https://www.danieldsjoberg.com/gtsummary/reference/modify.html) | update spanning headers |
| [`modify_caption()`](https://www.danieldsjoberg.com/gtsummary/reference/modify_caption.html) | update table caption/title |
| [`bold_labels()`](https://www.danieldsjoberg.com/gtsummary/reference/bold_italicize_labels_levels.html) | bold variable labels |
| [`bold_levels()`](https://www.danieldsjoberg.com/gtsummary/reference/bold_italicize_labels_levels.html) | bold variable levels |
| [`italicize_labels()`](https://www.danieldsjoberg.com/gtsummary/reference/bold_italicize_labels_levels.html) | italicize variable labels |
| [`italicize_levels()`](https://www.danieldsjoberg.com/gtsummary/reference/bold_italicize_labels_levels.html) | italicize variable levels |
| [`bold_p()`](https://www.danieldsjoberg.com/gtsummary/reference/bold_p.html) | bold significant p-values |

### Example 4: Multi-line Continuous Summaries

```{r}
trial |>
  tbl_summary(
    by = trt,
    include = age,
    type = all_continuous() ~ "continuous2",
    statistic = all_continuous() ~ c(
      "{N_nonmiss}",
      "{median} ({p25}, {p75})",
      "{min}, {max}"
    ),
    missing = "no"
  ) |>
  add_p(pvalue_fun = label_style_pvalue(digits = 2))
```

> To update the continuous variables to summarize on multiple lines, update the summary type to `"continuous2"` (for summaries on two or more lines).


### Example 5: Use `gt` functions to format table

```{r}
trial |>
  tbl_summary(by = trt, include = c(age, grade), missing = "no") |>
  add_n() |>
  as_gt() |>
  gt::tab_source_note(gt::md("*This data is simulated*"))
```

