---
title: "rpact"
search: true
---

```{r warning = F, message = F}
knitr::opts_chunk$set(
  dev = "png",          
  dpi = 300,
  fig.width = 7,
  fig.height = 5
)

library(tidyverse)
library(rpact)
library(PWEXP)

```

## Links {-}

+ [rpact website](https://www.rpact.org/#r-generics)
+ [rpact start guide](https://manual.rpact.org/pdf/rpact_v4_2.1_quick_start_guide.pdf)
+ [Step-by-step rpact tutorial](https://www.rpact.org/vignettes/getting-started/rpact_step-by-step_tutorial/)

## Start {-}

```{r}
# create an inverse normal design with default parameters
design <- getDesignInverseNormal()

design %>% summary()
```


```{r}
# display the design characteristics
design %>% getDesignCharacteristics()
```

```{r}
# plot the design with default type 1 (Boundary Plot)
design %>% plot()
```
## Example: PFS simulation using protocol assumptions {-}

This is an example for simulating the PFS events to get the predicted projection time and confidence internval based on the protocol assumption, considering actual enrollment and drop out rate per month. 

```{r}
ii <- 1:100

mon.enr.6 <- rep(40, 20)
start_date <- as.Date("2022-08-01")

get.followup.sap <- function(ii){
    set.seed(92480+ii)
    dat.sap.sim <- 
      simdata(group=c('trt','ctrl'), 
              allocation = c(1, 1), 
              n_rand = mon.enr.6, ## real enrollment
              event_lambda = c(log(2)/20*0.75,
                               log(2)/20),
              drop_rate = 0.008,  ## monthly rate
              total_sample = 800,
              add_column = c('followT','event', 'followT_abs','censor'))
    enr2 <- dat.sap.sim %>% 
      filter(event == 1) %>% 
      arrange(followT_abs) 
    return(enr2 %>% select(followT_abs) %>% 
             slice(c(344, 458)))
}
 
suppressMessages(
  sim.res.sap <- map_dfc(.x = ii, .f = get.followup.sap)
)

sim.res.sap.df <- as.data.frame(sim.res.sap)

```

```{r}
# interim analysis 
x344 <- quantile((as.numeric(sim.res.sap.df[1, ])), 
                 probs = c(0.025, 0.5, 0.975))
x344
start_date + x344*30.4375
```

```{r}
# final analysis 
x458 <- quantile((as.numeric(sim.res.sap.df[2, ])), 
                 probs = c(0.025, 0.5, 0.975))
x458
start_date+ x458*30.4375
```

## Example: event prediction based on the actual data {-}

#### Simulate PFS data {-}

Suppuse the actual follow up time is 36 months. 

```{r}
set.seed(12345)

mon.enr.6 <- rep(40, 20)

data <- dat.sap.sim <- 
      simdata(group=c('trt','ctrl'), 
              allocation = c(1, 1), 
              n_rand = mon.enr.6, ## real enrollment
              event_lambda = c(log(2)/25*0.5,
                               log(2)/25),
              drop_rate = 0.0065,  ## monthly rate
              total_sample = 800,
              add_column = c('followT','event', 'followT_abs','censor')) %>% 
  mutate(drop = as.numeric((followT_abs <= 36) & (censor == 1))) %>% 
  mutate(censor = as.numeric(drop == 1 | followT_abs > 36)) %>% 
  mutate(followT_abs = pmin(36, followT_abs), 
         followT = pmin(followT, 36 - randT)) %>% 
  mutate(event = 1 - censor)
```

```{r}
head(data)
```

```{r}
table(data$event)

table(data$drop)
```


```{r}
fit_b1 <- pwexp.fit(data$followT, data$event, nbreak = 1)  
fit_b1
```

```{r}
plot_survival(data$followT, data$drop,  
              ylab='Survival function of drop-out', 
              main='Fitted Model with 95% CI')
```


```{r warning = F, message = F}
invisible(capture.output(
  fit_b1_boot <- boot.pwexp.fit(fit_b1, nsim = 100, seed = 0102)
))

invisible(capture.output(
  fit_censor_boot1 <- boot.pwexp.fit(data$followT, 
                                   data$drop, nbreak = 1, 
                                   nsim = 100, seed=0102)
))

cut_indicator = 1- data$event
cut_indicator[data$drop==1]=0;

anat=(as.Date("2025-08-01") -
        as.Date("2022-08-01"))/30.4375

predicted_boot11 <- predict(fit_b1_boot, 
                            cut_indicator=cut_indicator,
                            analysis_time=anat,
                            censor_model_boot=fit_censor_boot1,
                            future_rand=NULL,seed=0102,
                            output = "summary" )
```

```{r}
plot.new()
pred_time11 <- plot_event(predicted_boot11, xyswitch=T, 
                          eval_at = c(344,458), alpha = 0.05)
pred_time11
```

```{r}
sumDt11 =as.data.frame(pred_time11); 
sumDt11$time =start_date + as.data.frame(pred_time11)$time*30.4375
sumDt11$`2.5% time` = start_date + as.data.frame(pred_time11)$`2.5% time`*30.4375; 
sumDt11$`97.5% time` = start_date + as.data.frame(pred_time11)$`97.5% time`*30.4375;

sumDt11
```

#### Event curve plot {-}

```{r}
accrualTime <- 
  getAccrualTime(accrualTime = 0:(20-1), 
                 accrualIntensity = rep(40, 20), 
                 maxNumberOfSubjects = 800)

y <- getEventProbabilities(time = 1:80, 
                           lambda2 = log(2)/20, 
                           accrualTime = accrualTime,
                           hazardRatio = 0.5, 
                           allocationRatioPlanned = 1s)
plot_obj <- rpact:::plot.EventProbabilities(y, plot = FALSE)
yValues <- plot_obj$data$yValues

plot_event(data$followT_abs, data$event, ylim = c(0, 500), 
           xlim = c(0, 80), col = "blue", main = "Event Curve", cex.axis = 0.65)
plot_event(predicted_boot11)

lines(x = 1:80, y = yValues[1:80], col = "grey30", lwd = 2)

legend("topleft", c("Actual", "CI", "Planned"), col = c("blue", "red", "grey30"),
       lwd = 2, cex = 0.65)

```

